Energy and exergy analysis of a cruise ship

### 2018-06-05 Manuscript/Manuscript.tex
Updated Appendix

\section{Energy system modelling}

@@

\subsection{Engine modelling, figures and tables} \label{sec:appendix:engines}

@@

\begin{table}
	\small
	\centering
	\begin{tabular}{llccrrrr}
		\hline
		\multicolumn{2}{l}{Variable} & Unit & M/C & 30\% & 50\% & 70\% & 90\% \\
		\hline
		\multicolumn{2}{l}{Compressor} & & & & & & \\ 
		& 	$T_{air,in}$  		& $^o$C & M & \multicolumn{4}{c}{302} \\ 
		&	$T_{air,out}$  		& $^o$C & C  & 356.0 & 409.4 & 434.1 & 461.7 \\ 
		&	$\dot{m}_{air}$  	& $kg/s$ & C  & 4.24 & 7.62 & 8.51 & 10.28 \\
		&   $\beta_{comp}$		& - 	& M & & & &  \\
		\multicolumn{2}{l}{Turbine} & & & & & &  \\
		&	$T_{eg,in}$   		& $^o$C & C  & 698.8 & 702.2 & 763.3 & 786.0 \\ 
		&	$T_{eg,out}$  		& $^o$C & M  & 639.3 & 592.9 & 633.2 & 631.0 \\ 
		&	$\dot{m}_{eg}$  	& $kg/s$ & C  & 4.35 & 7.79 & 8.73 & 10.57 \\
		&   $\beta_{exp}$		& - 	& C & & & &  \\
		\multicolumn{2}{l}{Bypass valve} & & & & & &  \\
		&	$\dot{m}_{air,BP}$	& $kg/s$ & C  & 0.98 & 2.05 & 1.02 & 0.39 \\  
		\multicolumn{2}{l}{Cylinders} & & & & & &   \\
		&	$\dot{m}_{air,in}$  & $kg/s$ & C  & 3.25 & 5.57 & 7.49 & 9.89 \\
		&	$\dot{m}_{fuel,in}$ & $kg/s$ & C  & 0.11 & 0.17 & 0.22 & 0.28 \\ 
		&	$\dot{m}_{eg,out}$  & $kg/s$ & C  & 3.37 & 5.74 & 7.71 & 10.17 \\
		&	$T_{air,in}$  		& $^o$C & M  & 323.7 & 323.2 & 323.1 & 323.5 \\ 
		&	$T_{eg,out}$  		& $^o$C & M  & 698.8 & 702.2 & 763.3 & 786.0 \\ 
		\multicolumn{2}{l}{CAC-LT stage} & & & & & &  \\
		&	$T_{air,in}$  		& $^o$C & C  & 355.2 & 350.1 & 360.0 & 368.7 \\ 
		&	$T_{air,out}$ 		& $^o$C & M  & 323.7 & 323.2 & 323.1 & 323.5 \\
		&	$T_{w,in}$   		& $^o$C & M  & 309.3 & 309.6 & 309.5 & 309.6 \\
		&	$T_{w,out}$   		& $^o$C & C  & 310.4 & 311.0 & 311.8 & 313.0 \\  
		& 	$\dot{m}_{w,LT}$  	& $kg/s$ & C& 23.36 & 26.09 & 29.1 & 31.53 \\
		\multicolumn{2}{l}{CAC-HT stage} & & & & & & \\
		&	$T_{air,in}$   		& $^o$C & C  & 356.0 & 409.4 & 434.1 & 461.7 \\
		&	$T_{air,out}$   	& $^o$C & C  & 355.2 & 350.1 & 360.0 & 368.7 \\ 
		&	$T_{w,in}$    		& $^o$C & C  & 360.8 & 360.2 & 359.6 & 358.7 \\ 
		&	$T_{w,out}$   		& $^o$C & C  & 360.8 & 362.6 & 363.6 & 365.4 \\
		& 	$\dot{m}_{w,HT}$  	& $kg/s$ & C & 33.21 & 33.18 & 33.33 & 33.33 \\
		\multicolumn{2}{l}{LOC} & & & & & &  \\
		&	$T_{lo,in}$   		& $^o$C & C  &346.6 & 348.9 & 349.8 & 350.5 \\ 
		&	$T_{lo,out}$    	& $^o$C & M  & 335.1 & 336.8 & 338.1 & 338.3 \\ 
		&	$T_{w,in}$    		& $^o$C & C  & 310.4 & 311.0 & 311.8 & 313.0 \\
		&	$T_{w,out}$   		& $^o$C & C  & 317.4 & 317.6 & 317.5 & 318.5 \\ 
		\multicolumn{2}{l}{JWC} & & & & & &  \\
		&	$T_{wall}$    		& $^o$C & C  & \multicolumn{4}{c}{423}  \\ 
		&	$T_{w,in}$    		& $^o$C & M  & 358.1 & 358.4 & 357.6 & 356.4 \\ 
		&	$T_{w,out}$	   		& $^o$C & C  & 360.8 & 360.2 & 359.6 & 358.7 \\ 
		\hline
	\end{tabular}
	\caption{Main engines, measured and calculated temperatures and flows at different engine loads}
	\label{tab:ME_values}
\end{table}

\begin{table}
	\small
	\centering
	\begin{tabular}{llccrrrr}
		\hline
		\multicolumn{2}{l}{Variable} & Unit & M/C & 30\% & 50\% & 70\% & 90\% \\
		\hline
		\multicolumn{2}{l}{Compressor} & & & & & & \\ 
		& 	$T_{air,in}$  		& $^o$C & M & 284.2 & 285.3 & 284.9 & 294.6 \\ 
		&	$T_{air,out}$  		& $^o$C & C  & 323.2 & 357.9 & 391.6 & 430.9 \\
		&	$\dot{m}_{air}$  	& $kg/s$ & C  & 1.96 & 2.65 & 3.5 & 4.27 \\
		&   $\beta_{comp}$		& - 	& M & & & &  \\
		\multicolumn{2}{l}{Turbine} & & & & & &  \\
		&	$T_{eg,in}$   		& $^o$C & M  & 729.5 & 762.1 & 770.3 & 798.5 \\
		&	$T_{eg,out}$  		& $^o$C & M  & 652.8 & 667.8 & 654.2 & 659.4 \\
		&	$\dot{m}_{eg}$  	& $kg/s$ & C  & 2.02 & 2.74 & 3.61 & 4.41 \\
		&   $\beta_{exp}$		& - 	& C & & & &  \\
		\multicolumn{2}{l}{Cylinders} & & & & & &   \\
		&	$\dot{m}_{air,in}$  & $kg/s$ & C  & 1.96 & 2.65 & 3.5 & 4.27 \\ 
		&	$\dot{m}_{fuel,in}$ & $kg/s$ & C  & 0.05 & 0.08 & 0.11 & 0.14 \\
		&	$\dot{m}_{eg,out}$  & $kg/s$ & C  & 2.02 & 2.74 & 3.61 & 4.41 \\
		&	$T_{air,in}$  		& $^o$C & M  & 320.8 & 322.3 & 327.6 & 329.1 \\
		&	$T_{eg,out}$  		& $^o$C & C  & 729.5 & 762.1 & 770.3 & 798.5 \\
		\multicolumn{2}{l}{CAC-LT stage} & & & & & &  \\
		&	$T_{air,in}$  		& $^o$C & C  & 323.2 & 356.3 & 366.7 & 372.0 \\ 
		&	$T_{air,out}$ 		& $^o$C & M  & 320.8 & 322.3 & 327.6 & 329.1 \\ 
		&	$T_{w,in}$   		& $^o$C & M  & 317.1 & 316.3 & 317.5 & 314.1 \\
		&	$T_{w,out}$   		& $^o$C & C  & 317.2 & 318.2 & 320.5 & 318.3 \\
		& 	$\dot{m}_{w,LT}$  	& $kg/s$ & C & 11.16 & 11.16 & 10.98 & 10.78 \\ 
		\multicolumn{2}{l}{CAC-HT stage} & & & & & & \\
		&	$T_{air,in}$   		& $^o$C & C  & 323.2 & 357.9 & 391.6 & 430.9 \\ 
		&	$T_{air,out}$   	& $^o$C & C  & 323.2 & 356.3 & 366.7 & 372.0 \\ 
		&	$T_{w,in}$    		& $^o$C & C  & 362.5 & 364.2 & 368.0 & 368.2 \\
		&	$T_{w,out}$   		& $^o$C & C  & 362.5 & 364.3 & 369.3 & 371.9 \\
		& 	$\dot{m}_{w,HT}$  	& $kg/s$ & C & 16.66 & 16.67 & 16.55 & 16.67 \\
		\multicolumn{2}{l}{LOC} & & & & & &  \\
		&	$T_{lo,in}$   		& $^o$C & C  & 348.2 & 350.3 & 354.0 & 357.3 \\
		&	$T_{lo,out}$    	& $^o$C & M  & 336.6 & 337.4 & 339.7 & 340.4 \\
		&	$T_{w,in}$    		& $^o$C & C & 317.2 & 318.2 & 320.5 & 318.3 \\ 
		&	$T_{w,out}$   		& $^o$C & C  & 325.9 & 327.8 & 331.4 & 331.3 \\
		\multicolumn{2}{l}{JWC} & & & & & & \\
		&	$T_{wall}$    		& $^o$C & C  & \multicolumn{4}{c}{423}  \\
		&	$T_{w,in}$    		& $^o$C & M  & 360.0 & 359.7 & 362.3 & 361.5 \\
		&	$T_{w,out}$	   		& $^o$C & C  & 362.5 & 364.2 & 368.0 & 368.2 \\  
		\hline
	\end{tabular}
	\caption{Auxiliary engines, measured and calculated temperatures and flows at different engine loads}
	\label{tab:AE_values}
\end{table}

\begin{table}
	\small
	\centering
	\begin{tabular}{lcrrrr}
		\hline
		Energy flow & type & 30\% & 50\% & 70\% & 90\% \\
		\hline
		Power output & Mech &  1767 & 2946 & 4118 & 5293 \\	
		Exhaust gas (after turbine) & Heat & 1672 & 2566 & 3227 & 3938 \\
		CAC-LT & Heat & 80 & 197 & 353 & 613 \\
		CAC-HT & Heat & 0 & 256  & 470 & 763  \\
		JWC & Heat & 501 & 507 & 549 & 620 \\
		LOC & Heat & 501 & 507 & 549 & 620 \\
		\hline
	\end{tabular}
	\caption{Main engines, energy flows at different engine loads}
	\label{tab:ME_flows}
\end{table}

\begin{table}
	\small
	\centering
	\begin{tabular}{lcrrrr}
		\hline
		Energy flow & type & 30\% & 50\% & 70\% & 90\% \\
		\hline
		Power output & Mech & 828 & 1381 & 1932 & 2482 \\

		Exhaust gas (after turbine) & Heat & 808 & 1144 & 1456 & 1753 \\
		CAC-LT & Heat & 5 & 92 & 139 & 185 \\
		CAC-HT & Heat & 0 & 4 & 87 & 256  \\
		JWC & Heat & 278 & 366 & 436 & 510 \\ 
		LOC & Heat & 278 & 366 & 436 & 510 \\ 
		\hline
	\end{tabular}
	\caption{Auxiliary engines, energy flows at different engine loads}
	\label{tab:AE_flows}
\end{table}

@@

\subsection{Electric demand modelling}

@@

Hence, the HVAC electric power demand was estimated as follows:
\begin{equation}
P_{el,HVAC} =
\begin{cases}
P_{el,tot}(t) - P_{el,ref}(t) , & \text{if}\ \text{2014-07-03} < t < \text{2014-08-21} \\
0, & \text{otherwise}
\end{cases}
\end{equation}
where $P_{el,ref}(t)$ is calculated as:
\begin{equation}
P_{el,ref}(t) = 0.5 (P_{el,tot}(t_1) + P_{el,tot}(t_2))
\end{equation}

@@

\subsubsection{Parameter estimation and uncertainty quantification}

@@

@@

@@

>>>>

\section{Extended method description}

In this appendix, a thorough description of the methods of this paper is provided, with particular focus on the data processing from raw measurements and ship technical documentation to the values required to calculate energy and exergy flows. These details were omitted in the main part of the paper to improve itse readability.

@@

\subsection{Diesel engines modeling} \label{sec:appendix:engines}

@@

\begin{table}
	\small
	\centering
	\begin{tabular}{llccrrrr}
		\toprule
		\multicolumn{2}{l}{Variable} & Unit & M/C & 30\% & 50\% & 70\% & 90\% \\
		\midrule
		\multicolumn{2}{l}{Compressor} & & & & & & \\ 
		& 	$T_{air,in}$  		& $^o$C & M & \multicolumn{4}{c}{302} \\ 
		&	$T_{air,out}$  		& $^o$C & C  & 356.0 & 409.4 & 434.1 & 461.7 \\ 
		&	$\dot{m}_{air}$  	& $kg/s$ & C  & 4.24 & 7.62 & 8.51 & 10.28 \\
		&   $\beta_{comp}$		& - 	& M & & & &  \\
		\multicolumn{2}{l}{Turbine} & & & & & &  \\
		&	$T_{eg,in}$   		& $^o$C & C  & 698.8 & 702.2 & 763.3 & 786.0 \\ 
		&	$T_{eg,out}$  		& $^o$C & M  & 639.3 & 592.9 & 633.2 & 631.0 \\ 
		&	$\dot{m}_{eg}$  	& $kg/s$ & C  & 4.35 & 7.79 & 8.73 & 10.57 \\
		&   $\beta_{exp}$		& - 	& C & & & &  \\
		\multicolumn{2}{l}{Bypass valve} & & & & & &  \\
		&	$\dot{m}_{air,BP}$	& $kg/s$ & C  & 0.98 & 2.05 & 1.02 & 0.39 \\  
		\multicolumn{2}{l}{Cylinders} & & & & & &   \\
		&	$\dot{m}_{air,in}$  & $kg/s$ & C  & 3.25 & 5.57 & 7.49 & 9.89 \\
		&	$\dot{m}_{fuel,in}$ & $kg/s$ & C  & 0.11 & 0.17 & 0.22 & 0.28 \\ 
		&	$\dot{m}_{eg,out}$  & $kg/s$ & C  & 3.37 & 5.74 & 7.71 & 10.17 \\
		&	$T_{air,in}$  		& $^o$C & M  & 323.7 & 323.2 & 323.1 & 323.5 \\ 
		&	$T_{eg,out}$  		& $^o$C & M  & 698.8 & 702.2 & 763.3 & 786.0 \\ 
		\multicolumn{2}{l}{CAC-LT stage} & & & & & &  \\
		&	$T_{air,in}$  		& $^o$C & C  & 355.2 & 350.1 & 360.0 & 368.7 \\ 
		&	$T_{air,out}$ 		& $^o$C & M  & 323.7 & 323.2 & 323.1 & 323.5 \\
		&	$T_{w,in}$   		& $^o$C & M  & 309.3 & 309.6 & 309.5 & 309.6 \\
		&	$T_{w,out}$   		& $^o$C & C  & 310.4 & 311.0 & 311.8 & 313.0 \\  
		& 	$\dot{m}_{w,LT}$  	& $kg/s$ & C& 23.36 & 26.09 & 29.1 & 31.53 \\
		\multicolumn{2}{l}{CAC-HT stage} & & & & & & \\
		&	$T_{air,in}$   		& $^o$C & C  & 356.0 & 409.4 & 434.1 & 461.7 \\
		&	$T_{air,out}$   	& $^o$C & C  & 355.2 & 350.1 & 360.0 & 368.7 \\ 
		&	$T_{w,in}$    		& $^o$C & C  & 360.8 & 360.2 & 359.6 & 358.7 \\ 
		&	$T_{w,out}$   		& $^o$C & C  & 360.8 & 362.6 & 363.6 & 365.4 \\
		& 	$\dot{m}_{w,HT}$  	& $kg/s$ & C & 33.21 & 33.18 & 33.33 & 33.33 \\
		\multicolumn{2}{l}{LOC} & & & & & &  \\
		&	$T_{lo,in}$   		& $^o$C & C  &346.6 & 348.9 & 349.8 & 350.5 \\ 
		&	$T_{lo,out}$    	& $^o$C & M  & 335.1 & 336.8 & 338.1 & 338.3 \\ 
		&	$T_{w,in}$    		& $^o$C & C  & 310.4 & 311.0 & 311.8 & 313.0 \\
		&	$T_{w,out}$   		& $^o$C & C  & 317.4 & 317.6 & 317.5 & 318.5 \\ 
		\multicolumn{2}{l}{JWC} & & & & & &  \\
		&	$T_{wall}$    		& $^o$C & C  & \multicolumn{4}{c}{423}  \\ 
		&	$T_{w,in}$    		& $^o$C & M  & 358.1 & 358.4 & 357.6 & 356.4 \\ 
		&	$T_{w,out}$	   		& $^o$C & C  & 360.8 & 360.2 & 359.6 & 358.7 \\ 
		\bottomrule
	\end{tabular}
	\caption{Main engines, measured and calculated temperatures and flows at different engine loads}
	\label{tab:ME_values}
\end{table}

\begin{table}
	\small
	\centering
	\begin{tabular}{llccrrrr}
		\toprule
		\multicolumn{2}{l}{Variable} & Unit & M/C & 30\% & 50\% & 70\% & 90\% \\
		\midrule
		\multicolumn{2}{l}{Compressor} & & & & & & \\ 
		& 	$T_{air,in}$  		& $^o$C & M & 284.2 & 285.3 & 284.9 & 294.6 \\ 
		&	$T_{air,out}$  		& $^o$C & C  & 323.2 & 357.9 & 391.6 & 430.9 \\
		&	$\dot{m}_{air}$  	& $kg/s$ & C  & 1.96 & 2.65 & 3.5 & 4.27 \\
		&   $\beta_{comp}$		& - 	& M & & & &  \\
		\multicolumn{2}{l}{Turbine} & & & & & &  \\
		&	$T_{eg,in}$   		& $^o$C & M  & 729.5 & 762.1 & 770.3 & 798.5 \\
		&	$T_{eg,out}$  		& $^o$C & M  & 652.8 & 667.8 & 654.2 & 659.4 \\
		&	$\dot{m}_{eg}$  	& $kg/s$ & C  & 2.02 & 2.74 & 3.61 & 4.41 \\
		&   $\beta_{exp}$		& - 	& C & & & &  \\
		\multicolumn{2}{l}{Cylinders} & & & & & &   \\
		&	$\dot{m}_{air,in}$  & $kg/s$ & C  & 1.96 & 2.65 & 3.5 & 4.27 \\ 
		&	$\dot{m}_{fuel,in}$ & $kg/s$ & C  & 0.05 & 0.08 & 0.11 & 0.14 \\
		&	$\dot{m}_{eg,out}$  & $kg/s$ & C  & 2.02 & 2.74 & 3.61 & 4.41 \\
		&	$T_{air,in}$  		& $^o$C & M  & 320.8 & 322.3 & 327.6 & 329.1 \\
		&	$T_{eg,out}$  		& $^o$C & C  & 729.5 & 762.1 & 770.3 & 798.5 \\
		\multicolumn{2}{l}{CAC-LT stage} & & & & & &  \\
		&	$T_{air,in}$  		& $^o$C & C  & 323.2 & 356.3 & 366.7 & 372.0 \\ 
		&	$T_{air,out}$ 		& $^o$C & M  & 320.8 & 322.3 & 327.6 & 329.1 \\ 
		&	$T_{w,in}$   		& $^o$C & M  & 317.1 & 316.3 & 317.5 & 314.1 \\
		&	$T_{w,out}$   		& $^o$C & C  & 317.2 & 318.2 & 320.5 & 318.3 \\
		& 	$\dot{m}_{w,LT}$  	& $kg/s$ & C & 11.16 & 11.16 & 10.98 & 10.78 \\ 
		\multicolumn{2}{l}{CAC-HT stage} & & & & & & \\
		&	$T_{air,in}$   		& $^o$C & C  & 323.2 & 357.9 & 391.6 & 430.9 \\ 
		&	$T_{air,out}$   	& $^o$C & C  & 323.2 & 356.3 & 366.7 & 372.0 \\ 
		&	$T_{w,in}$    		& $^o$C & C  & 362.5 & 364.2 & 368.0 & 368.2 \\
		&	$T_{w,out}$   		& $^o$C & C  & 362.5 & 364.3 & 369.3 & 371.9 \\
		& 	$\dot{m}_{w,HT}$  	& $kg/s$ & C & 16.66 & 16.67 & 16.55 & 16.67 \\
		\multicolumn{2}{l}{LOC} & & & & & &  \\
		&	$T_{lo,in}$   		& $^o$C & C  & 348.2 & 350.3 & 354.0 & 357.3 \\
		&	$T_{lo,out}$    	& $^o$C & M  & 336.6 & 337.4 & 339.7 & 340.4 \\
		&	$T_{w,in}$    		& $^o$C & C & 317.2 & 318.2 & 320.5 & 318.3 \\ 
		&	$T_{w,out}$   		& $^o$C & C  & 325.9 & 327.8 & 331.4 & 331.3 \\
		\multicolumn{2}{l}{JWC} & & & & & & \\
		&	$T_{wall}$    		& $^o$C & C  & \multicolumn{4}{c}{423}  \\
		&	$T_{w,in}$    		& $^o$C & M  & 360.0 & 359.7 & 362.3 & 361.5 \\
		&	$T_{w,out}$	   		& $^o$C & C  & 362.5 & 364.2 & 368.0 & 368.2 \\  
		\bottomrule
	\end{tabular}
	\caption{Auxiliary engines, measured and calculated temperatures and flows at different engine loads}
	\label{tab:AE_values}
\end{table}

\begin{table}
	\small
	\centering
	\begin{tabular}{lcrrrr}
		\toprule
		Energy flow & type & 30\% & 50\% & 70\% & 90\% \\
		\midrule
		Power output & Mech &  1767 & 2946 & 4118 & 5293 \\	
		Exhaust gas (after turbine) & Heat & 1672 & 2566 & 3227 & 3938 \\
		CAC-LT & Heat & 80 & 197 & 353 & 613 \\
		CAC-HT & Heat & 0 & 256  & 470 & 763  \\
		JWC & Heat & 501 & 507 & 549 & 620 \\
		LOC & Heat & 501 & 507 & 549 & 620 \\
		\bottomrule
	\end{tabular}
	\caption{Main engines, energy flows at different engine loads}
	\label{tab:ME_flows}
\end{table}

\begin{table}
	\small
	\centering
	\begin{tabular}{lcrrrr}
		\toprule
		Energy flow & type & 30\% & 50\% & 70\% & 90\% \\
		\midrule
		Power output & Mech & 828 & 1381 & 1932 & 2482 \\
		Exhaust gas (after turbine) & Heat & 808 & 1144 & 1456 & 1753 \\
		CAC-LT & Heat & 5 & 92 & 139 & 185 \\
		CAC-HT & Heat & 0 & 4 & 87 & 256  \\
		JWC & Heat & 278 & 366 & 436 & 510 \\ 
		LOC & Heat & 278 & 366 & 436 & 510 \\ 
		\bottomrule
	\end{tabular}
	\caption{Auxiliary engines, energy flows at different engine loads}
	\label{tab:AE_flows}
\end{table}

@@

\subsection{Electric demand modeling}

@@

The HVAC electric power demand was estimated as follows:
\begin{equation}
P_{el,HVAC} =
\begin{cases}
P_{el,tot}(t) - P_{el,ref}(t) , & \text{if}\ \text{2014-07-03} < t < \text{2014-08-21} \\
0, & \text{otherwise}
\end{cases}
\end{equation}
where $P_{el,ref}(t)$ is calculated as:
\begin{equation}
P_{el,ref}(t) = 0.5 (P_{el,tot}(t_1) + P_{el,tot}(t_2))
\end{equation}

@@

\subsubsection{Uncertainty quantification of heat demand}

@@

In this appendix, the cumulated yearly values for the main ship energy and exergy flows are provided. Apart from yearly values, we also provide the same statistics separated based on the ship operational mode (P/S: Port stay, LSS: Low speed sailing, Man: Maneuvering, HSS, High speed sailing) and based on the season (Win: Winter, Mid: Spring/autumn, Sum: Summer).

@@

In this section, the full list of all measurements available in the dataset that was used in this paper is provided. It should be noted that not all the values were used in the analysis.

@@

\section{Energy demand for typical operational days}

In this section, the values for the main energy demands (Propulsion power, electric power, high temperature heat and low temperature heat) are provided for the typical days identified in the demand clustering.
